---
title: "Untitled"
author: "Salma"
date: "2026-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(caret)
library(glmnet)
load("/Users/salma/Library/CloudStorage/OneDrive-UniversityofArizona/Desktop/Thesis_grad/scan_MNLR.RData")
head(scan)
```


```{r }


# ── STEP 2: 80:20 split ───────────────────────────────────────────
set.seed(123)
trainIndex_80 <- createDataPartition(
  scan$Behaviour_Coarse, p = 0.80, list = FALSE)
train_80 <- scan[ trainIndex_80, ]
test_20  <- scan[-trainIndex_80, ]

```

```{r}

# ── STEP 3: Factor reference levels ──────────────────────────────
train_80$Behaviour_Coarse <- relevel(
  as.factor(train_80$Behaviour_Coarse), ref = "Rest")
train_80$Age.Class <- relevel(
  as.factor(train_80$Age.Class), ref = "Adult")
train_80$Sex <- relevel(
  as.factor(train_80$Sex), ref = "Female")

test_20$Behaviour_Coarse <- factor(
  test_20$Behaviour_Coarse, levels = levels(train_80$Behaviour_Coarse))
test_20$Age.Class <- factor(
  test_20$Age.Class, levels = levels(train_80$Age.Class))
test_20$Sex <- factor(
  test_20$Sex, levels = levels(train_80$Sex))
```


```{r}

# ── STEP 4: CRITICAL — set location levels CORRECTLY ─────────────
# Do this BEFORE model.matrix — this is the key fix
train_80$Location_Coarse <- factor(
  train_80$Location_Coarse,
  levels = c("Forest", "Roadside", "RoadsideVeg")
)
test_20$Location_Coarse <- factor(
  test_20$Location_Coarse,
  levels = c("Forest", "Roadside", "RoadsideVeg")
)

# Verify
cat("Location levels (must be Forest, Roadside, RoadsideVeg):\n")
print(levels(train_80$Location_Coarse))
```


```{r}

# ── STEP 5: Circular time encoding ───────────────────────────────
train_80$hour     <- as.numeric(format(train_80$POSIX, "%H"))
train_80$hour_sin <- sin(2 * pi * train_80$hour / 24)
train_80$hour_cos <- cos(2 * pi * train_80$hour / 24)

test_20$hour      <- as.numeric(format(test_20$POSIX, "%H"))
test_20$hour_sin  <- sin(2 * pi * test_20$hour / 24)
test_20$hour_cos  <- cos(2 * pi * test_20$hour / 24)

```


```{r}

# ── STEP 6: Build model matrix ───────────────────────────────────
X_train_80 <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
  data = train_80)[, -1]

X_test_20 <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
  data = test_20)[, -1]

y_train_80 <- train_80$Behaviour_Coarse
y_test_20  <- test_20$Behaviour_Coarse

cat("Matrix dimensions:\n")
cat("Train:", nrow(X_train_80), "x", ncol(X_train_80), "\n")
cat("Test: ", nrow(X_test_20),  "x", ncol(X_test_20),  "\n")
```

```{r}

# ── STEP 7: Refit LASSO 
set.seed(123)
lasso_final_80 <- cv.glmnet(
  x      = X_train_80,
  y      = y_train_80,
  family = "multinomial",
  alpha  = 1,
  nfolds = 10
)

cat("LASSO retrained. lambda.min =", lasso_final_80$lambda.min, "\n")
```

```{r}
# ── STEP 7: Count non-zero coefficients ──────────────────────────
beta    <- lasso_final_80$glmnet.fit$beta
lam_idx <- which.min(abs(
  lasso_final_80$glmnet.fit$lambda - lasso_final_80$lambda.min))

cat("\nNon-zero coefficients at lambda.min:\n")
total <- 0
for (nm in names(beta)) {
  nz <- sum(beta[[nm]][, lam_idx] != 0)
  cat(sprintf("  %-20s: %d\n", nm, nz))
  total <- total + nz
}
cat(sprintf("\nTotal non-zero: %d\n", total))
cat(sprintf("Total possible: %d (36 predictors x 4 classes)\n", 36*4))
```

```{r}

# ── STEP 8: Verify HD probabilities are now correct ───────────────
hour_fixed <- 8
test_locs <- data.frame(
  Location_Coarse = factor(
    c("Forest", "Roadside", "RoadsideVeg"),
    levels = c("Forest", "Roadside", "RoadsideVeg")
  ),
  Age.Class = factor("Adult",  levels = levels(train_80$Age.Class)),
  Sex       = factor("Female", levels = levels(train_80$Sex)),
  hour      = hour_fixed,
  hour_sin  = sin(2 * pi * hour_fixed / 24),
  hour_cos  = cos(2 * pi * hour_fixed / 24)
)

X_check <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
  data = test_locs)[, -1]

prob_check <- predict(lasso_final_80,
                      newx = X_check,
                      s    = "lambda.min",
                      type = "response")[,,1]
colnames(prob_check) <- levels(y_train_80)
rownames(prob_check) <- c("Forest", "Roadside", "RoadsideVeg")

cat("\nHuman Directed probabilities (Roadside must be highest):\n")
print(round(prob_check[, "Human Directed", drop = FALSE], 3))

```
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
# Reextract model after retraining
lasso_fit   <- lasso_final_80$glmnet.fit
best_lambda <- lasso_final_80$lambda.min

# Run curves only first to verify
fig1a <- make_fig1_curves("Adult",    "Female", "fig1a")
fig1b <- make_fig1_curves("Adult",    "Male",   "fig1b")
fig1c <- make_fig1_curves("Subadult", "Female", "fig1c")
fig1d <- make_fig1_curves("Subadult", "Male",   "fig1d")
```
```{r}
getwd()


# Then move the file there in Finder, or set working directory to where the file is:
setwd("/Users/salma/Library/CloudStorage/OneDrive-UniversityofArizona/Desktop/Thesis_grad/")

# Then source works without full path
source("fig1_bootstrap_ci.R")
```


```{r}
# ── Set N_BOOT low for quick test first ──────────────────────────
N_BOOT     <- 50   # change to 500 for final run
hours_fine <- seq(8, 18, by = 0.5)
beh_levels <- levels(y_train_80)
margin     <- ggplot2::margin

loc_colors <- c(
  "Forest"      = "#1D9E75",
  "Roadside"    = "#E24B4A",
  "RoadsideVeg" = "#7F77DD"
)

lasso_fit   <- lasso_final_80$glmnet.fit
best_lambda <- lasso_final_80$lambda.min

make_time_grid <- function(age, sex, loc) {
  data.frame(
    Location_Coarse = factor(loc,
                             levels = c("Forest","Roadside","RoadsideVeg")),
    Age.Class       = factor(age,  levels = levels(train_80$Age.Class)),
    Sex             = factor(sex,  levels = levels(train_80$Sex)),
    hour            = hours_fine,
    hour_sin        = sin(2 * pi * hours_fine / 24),
    hour_cos        = cos(2 * pi * hours_fine / 24)
  )
}

get_probs <- function(grid, fit = lasso_fit, lam = best_lambda) {
  X_new <- model.matrix(
    ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
    data = grid
  )[, -1]
  prob <- predict(fit, newx = X_new, s = lam,
                  type = "response")[,,1]
  colnames(prob) <- beh_levels
  df          <- as.data.frame(prob)
  df$Hour     <- grid$hour
  df$Location <- as.character(grid$Location_Coarse[1])
  df
}

get_point_estimates <- function(age, sex) {
  bind_rows(lapply(
    c("Forest", "Roadside", "RoadsideVeg"),
    function(loc) get_probs(make_time_grid(age, sex, loc))
  ))
}

obs_props <- function(age, sex) {
  totals <- train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour) |>
    summarise(n_total = n(), .groups = "drop")
  train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour, Behaviour_Coarse) |>
    summarise(n_beh = n(), .groups = "drop") |>
    left_join(totals, by = c("Location_Coarse", "hour")) |>
    mutate(prop = n_beh / n_total) |>
    filter(n_total >= 2) |>
    rename(Behaviour = Behaviour_Coarse,
           Hour      = hour,
           Location  = Location_Coarse)
}

boot_ci_time <- function(age, sex, n_boot = N_BOOT) {
  locs    <- c("Forest", "Roadside", "RoadsideVeg")
  results <- list()
  for (loc in locs) {
    sub   <- train_80 |>
      filter(Age.Class == age, Sex == sex,
             Location_Coarse == loc)
    n_sub <- nrow(sub)
    if (n_sub < 3) {
      cat(sprintf("  Skipping %s %s %s — n=%d\n", age, sex, loc, n_sub))
      next
    }
    n_hours  <- length(hours_fine)
    boot_mat <- array(NA,
                      dim      = c(n_boot, n_hours, length(beh_levels)),
                      dimnames = list(NULL, NULL, beh_levels))
    n_success <- 0
    for (b in seq_len(n_boot)) {
      idx_boot   <- sample(seq_len(nrow(train_80)), replace = TRUE)
      train_boot <- train_80[idx_boot, ]
      # CRITICAL: re-enforce factor levels after resampling
      train_boot$Location_Coarse <- factor(
        train_boot$Location_Coarse,
        levels = c("Forest", "Roadside", "RoadsideVeg"))
      train_boot$Age.Class <- factor(
        train_boot$Age.Class,
        levels = levels(train_80$Age.Class))
      train_boot$Sex <- factor(
        train_boot$Sex,
        levels = levels(train_80$Sex))
      train_boot$Behaviour_Coarse <- factor(
        train_boot$Behaviour_Coarse,
        levels = levels(train_80$Behaviour_Coarse))
      X_boot <- tryCatch(
        model.matrix(
          ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
          data = train_boot)[, -1],
        error = function(e) NULL)
      if (is.null(X_boot)) next
      y_boot <- train_boot$Behaviour_Coarse
      fit_boot <- tryCatch(
        cv.glmnet(x = X_boot, y = y_boot,
                  family = "multinomial", alpha = 1, nfolds = 5),
        error = function(e) NULL)
      if (is.null(fit_boot)) next
      grid <- make_time_grid(age, sex, loc)
      prob <- tryCatch(
        get_probs(grid,
                  fit = fit_boot$glmnet.fit,
                  lam = fit_boot$lambda.min),
        error = function(e) NULL)
      if (is.null(prob)) next
      boot_mat[b, , ] <- as.matrix(prob[, beh_levels])
      n_success <- n_success + 1
    }
    cat(sprintf("  Bootstrap done: %s %s %s (n=%d, %d/%d ok)\n",
                age, sex, loc, n_sub, n_success, n_boot))
    ci_list <- lapply(seq_along(beh_levels), function(k) {
      vals <- boot_mat[, , k]
      vals <- vals[complete.cases(vals), , drop = FALSE]
      data.frame(
        Hour      = hours_fine,
        Behaviour = beh_levels[k],
        Lower     = apply(vals, 2, quantile, 0.025, na.rm = TRUE),
        Upper     = apply(vals, 2, quantile, 0.975, na.rm = TRUE),
        Location  = loc)
    })
    results[[loc]] <- bind_rows(ci_list)
  }
  bind_rows(results)
}

make_fig1_boot <- function(age, sex, fig_label) {
  cat(sprintf("\nBuilding %s (%s %s)...\n", fig_label, age, sex))
  fig_letter <- substr(fig_label, nchar(fig_label), nchar(fig_label))
  point_df <- get_point_estimates(age, sex) |>
    pivot_longer(cols = all_of(beh_levels),
                 names_to = "Behaviour", values_to = "Prob") |>
    mutate(Behaviour = factor(Behaviour, levels = beh_levels),
           Location  = factor(Location,
                              levels = c("Forest","Roadside","RoadsideVeg")))
  ci_df <- boot_ci_time(age, sex) |>
    mutate(Behaviour = factor(Behaviour, levels = beh_levels),
           Location  = factor(Location,
                              levels = c("Forest","Roadside","RoadsideVeg")))
  ribbon_df <- ci_df |>
    left_join(point_df, by = c("Hour", "Behaviour", "Location"))
  obs_df <- obs_props(age, sex) |>
    mutate(Behaviour = factor(Behaviour, levels = beh_levels),
           Location  = factor(Location,
                              levels = c("Forest","Roadside","RoadsideVeg")))
  p <- ggplot() +
geom_errorbar(
  data  = ribbon_df,
  aes(x    = Hour,
      ymin = Lower,
      ymax = Upper,
      color = Location),
  width     = 0.3,    # width of the horizontal caps
  linewidth = 0.5,
  alpha     = 0.7
) +
    geom_line(data = point_df,
              aes(x = Hour, y = Prob, color = Location),
              linewidth = 1.1) +
    geom_point(data = obs_df,
               aes(x = Hour, y = prop, color = Location),
               size = 2.5, shape = 21, fill = "white",
               stroke = 1.2, alpha = 0.85) +
    facet_wrap(~ Behaviour, nrow = 1, scales = "free_y") +
    scale_color_manual(values = loc_colors, name = "Location") +
    scale_x_continuous(
      limits = c(8, 18),
      breaks = c(8, 12, 18),
      labels = c("08:00","12:00","18:00")) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = expansion(mult = c(0.02, 0.10))) +
    labs(
      title    = sprintf(
        "Fig. 1%s \u2014 Predicted probability of behaviour across time of day (%s %s)",
        fig_letter, age, sex),
      subtitle = paste0("Age = ", age, " \u00b7 Sex = ", sex,
        " \u00b7 curves = LASSO predictions",
        " \u00b7 shading = bootstrap 95% CI (n=", N_BOOT, ")",
        " \u00b7 circles = observed proportions (n \u2265 2)"),
      x       = "Hour of day",
      y       = "Probability of Behaviour",
      caption = "Unweighted LASSO \u00b7 \u03bb = lambda.min \u00b7 Green = Forest \u00b7 Red = Roadside \u00b7 Purple = RoadsideVeg") +
    theme_minimal(base_size = 11) +
    theme(
      plot.background  = element_rect(fill = "white", color = NA),
      panel.background = element_rect(fill = "white", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
      strip.text       = element_text(size = 10, face = "bold", color = "grey15"),
      strip.background = element_rect(fill = "grey97", color = "grey88"),
      axis.text.x      = element_text(angle = 30, hjust = 1, size = 8.5),
      axis.text.y      = element_text(size = 9),
      legend.position  = "top",
      legend.title     = element_text(size = 9.5, face = "bold"),
      legend.text      = element_text(size = 9.5),
      plot.title       = element_text(size = 11, face = "bold", color = "grey10"),
      plot.subtitle    = element_text(size = 8.5, color = "grey45"),
      plot.caption     = element_text(size = 7.5, color = "grey55", hjust = 0),
      plot.margin      = unit(c(10, 10, 8, 10), "pt"))
  fname <- sprintf("fig1%s_%s_%s_boot.png",
                   fig_letter, tolower(age), tolower(sex))
  ggsave(fname, plot = p, width = 15, height = 6, dpi = 300, bg = "white")
  cat(sprintf("Saved: %s\n", fname))
  p
}

# ── Run all 4 figures (N_BOOT=50 for test, change to 500 for final)
set.seed(123)
fig1a_boot <- make_fig1_boot("Adult",    "Female", "fig1a")
fig1b_boot <- make_fig1_boot("Adult",    "Male",   "fig1b")
fig1c_boot <- make_fig1_boot("Subadult", "Female", "fig1c")
fig1d_boot <- make_fig1_boot("Subadult", "Male",   "fig1d")

cat("\nAll Fig 1 bootstrap plots complete.\n")
```

```{r}
# ══════════════════════════════════════════════════════════════════
# FIG 1 — Predicted probability across time of day
# Bootstrap 95% CI on LASSO curve by refitting on resamples
# Hours: 08:00–18:00
# All 4 Age×Sex combinations
# ══════════════════════════════════════════════════════════════════

library(ggplot2)
library(dplyr)
library(tidyr)
library(glmnet)

margin     <- ggplot2::margin
set.seed(123)

N_BOOT     <- 500   # set to 50 for quick test
hours_fine <- seq(8, 18, by = 0.5)
beh_levels <- levels(y_train_80)

loc_colors <- c(
  "Forest"      = "#1D9E75",
  "Roadside"    = "#E24B4A",
  "RoadsideVeg" = "#7F77DD"
)

# ── Extract fitted model ──────────────────────────────────────────
lasso_fit   <- lasso_final_80$glmnet.fit
best_lambda <- lasso_final_80$lambda.min

# ── Build prediction grid ─────────────────────────────────────────
make_time_grid <- function(age, sex, loc) {
  data.frame(
    Location_Coarse = factor(loc,
                             levels = c("Forest","Roadside","RoadsideVeg")),
    Age.Class       = factor(age,  levels = levels(train_80$Age.Class)),
    Sex             = factor(sex,  levels = levels(train_80$Sex)),
    hour            = hours_fine,
    hour_sin        = sin(2 * pi * hours_fine / 24),
    hour_cos        = cos(2 * pi * hours_fine / 24)
  )
}

# ── Get LASSO predicted probabilities ────────────────────────────
get_probs <- function(grid, fit = lasso_fit, lam = best_lambda) {
  X_new <- model.matrix(
    ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
    data = grid
  )[, -1]
  prob <- predict(fit,
                  newx = X_new,
                  s    = lam,
                  type = "response")[,,1]
  colnames(prob) <- beh_levels
  df          <- as.data.frame(prob)
  df$Hour     <- grid$hour
  df$Location <- as.character(grid$Location_Coarse[1])
  df
}

# ── Point estimates for all 3 locations ──────────────────────────
get_point_estimates <- function(age, sex) {
  bind_rows(lapply(
    c("Forest", "Roadside", "RoadsideVeg"),
    function(loc) get_probs(make_time_grid(age, sex, loc))
  ))
}

# ── Observed proportions (circles) ───────────────────────────────
obs_props <- function(age, sex) {
  totals <- train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour) |>
    summarise(n_total = n(), .groups = "drop")

  train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour, Behaviour_Coarse) |>
    summarise(n_beh = n(), .groups = "drop") |>
    left_join(totals, by = c("Location_Coarse", "hour")) |>
    mutate(prop = n_beh / n_total) |>
    filter(n_total >= 2) |>
    rename(Behaviour = Behaviour_Coarse,
           Hour      = hour,
           Location  = Location_Coarse)
}

# ── Bootstrap CI ─────────────────────────────────────────────────
# Resamples full training data, refits LASSO each time,
# predicts curve, gets 2.5/97.5 percentiles across resamples
# Factor levels enforced inside loop — prevents Roadside/RoadsideVeg swap

boot_ci_time <- function(age, sex, n_boot = N_BOOT) {

  locs    <- c("Forest", "Roadside", "RoadsideVeg")
  results <- list()

  for (loc in locs) {

    sub   <- train_80 |>
      filter(Age.Class == age,
             Sex        == sex,
             Location_Coarse == loc)

    n_sub <- nrow(sub)
    if (n_sub < 3) {
      cat(sprintf("  Skipping %s %s %s — n=%d\n",
                  age, sex, loc, n_sub))
      next
    }

    n_hours  <- length(hours_fine)
    boot_mat <- array(NA,
                      dim      = c(n_boot, n_hours, length(beh_levels)),
                      dimnames = list(NULL, NULL, beh_levels))

    n_success <- 0

    for (b in seq_len(n_boot)) {

      # Resample full training set with replacement
      idx_boot   <- sample(seq_len(nrow(train_80)), replace = TRUE)
      train_boot <- train_80[idx_boot, ]

      # CRITICAL: re-enforce correct factor levels after resampling
      # Without this resampling resets alphabetical order
      # which swaps Roadside and RoadsideVeg predictions
      train_boot$Location_Coarse <- factor(
        train_boot$Location_Coarse,
        levels = c("Forest", "Roadside", "RoadsideVeg"))
      train_boot$Age.Class <- factor(
        train_boot$Age.Class,
        levels = levels(train_80$Age.Class))
      train_boot$Sex <- factor(
        train_boot$Sex,
        levels = levels(train_80$Sex))
      train_boot$Behaviour_Coarse <- factor(
        train_boot$Behaviour_Coarse,
        levels = levels(train_80$Behaviour_Coarse))

      # Rebuild model matrix from bootstrap sample
      X_boot <- tryCatch(
        model.matrix(
          ~ (Location_Coarse + Age.Class + Sex +
             hour_sin + hour_cos)^3,
          data = train_boot)[, -1],
        error = function(e) NULL)
      if (is.null(X_boot)) next

      y_boot <- train_boot$Behaviour_Coarse

      # Refit LASSO on bootstrap sample
      fit_boot <- tryCatch(
        cv.glmnet(
          x      = X_boot,
          y      = y_boot,
          family = "multinomial",
          alpha  = 1,
          nfolds = 5),
        error = function(e) NULL)
      if (is.null(fit_boot)) next

      # Predict on fixed grid
      grid <- make_time_grid(age, sex, loc)
      prob <- tryCatch(
        get_probs(grid,
                  fit = fit_boot$glmnet.fit,
                  lam = fit_boot$lambda.min),
        error = function(e) NULL)
      if (is.null(prob)) next

      boot_mat[b, , ] <- as.matrix(prob[, beh_levels])
      n_success <- n_success + 1
    }

    cat(sprintf("  Bootstrap done: %s %s %s (n=%d, %d/%d ok)\n",
                age, sex, loc, n_sub, n_success, n_boot))

    # Compute 2.5th and 97.5th percentiles
    ci_list <- lapply(seq_along(beh_levels), function(k) {
      vals <- boot_mat[, , k]
      vals <- vals[complete.cases(vals), , drop = FALSE]
      data.frame(
        Hour      = hours_fine,
        Behaviour = beh_levels[k],
        Lower     = apply(vals, 2, quantile, 0.025, na.rm = TRUE),
        Upper     = apply(vals, 2, quantile, 0.975, na.rm = TRUE),
        Location  = loc)
    })

    results[[loc]] <- bind_rows(ci_list)
  }

  bind_rows(results)
}

# ── Main plot function ────────────────────────────────────────────
make_fig1_boot <- function(age, sex, fig_label) {

  cat(sprintf("\nBuilding %s (%s %s)...\n", fig_label, age, sex))

  fig_letter <- substr(fig_label, nchar(fig_label), nchar(fig_label))

  # Point estimates
  point_df <- get_point_estimates(age, sex) |>
    pivot_longer(cols      = all_of(beh_levels),
                 names_to  = "Behaviour",
                 values_to = "Prob") |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # Bootstrap CIs
  ci_df <- boot_ci_time(age, sex) |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # Merge CI with point estimates
  ribbon_df <- ci_df |>
    left_join(point_df, by = c("Hour", "Behaviour", "Location"))

  # Observed proportions
  obs_df <- obs_props(age, sex) |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # ── Plot ─────────────────────────────────────────────────────────
  p <- ggplot() +

    # Lower CI bound dashed line
    geom_line(
      data      = ribbon_df,
      aes(x = Hour, y = Lower, color = Location),
      linewidth = 0.4,
      linetype  = "dashed",
      alpha     = 0.6
    ) +

    # Upper CI bound dashed line
    geom_line(
      data      = ribbon_df,
      aes(x = Hour, y = Upper, color = Location),
      linewidth = 0.4,
      linetype  = "dashed",
      alpha     = 0.6
    ) +

    # LASSO point estimate — solid line
    geom_line(
      data      = point_df,
      aes(x = Hour, y = Prob, color = Location),
      linewidth = 1.1
    ) +

    # Observed proportion circles
    geom_point(
      data   = obs_df,
      aes(x = Hour, y = prop, color = Location),
      size   = 2.5,
      shape  = 21,
      fill   = "white",
      stroke = 1.2,
      alpha  = 0.85
    ) +

    facet_wrap(~ Behaviour, nrow = 1, scales = "free_y") +
    scale_color_manual(values = loc_colors, name = "Location") +
    scale_x_continuous(
      limits = c(8, 18),
      breaks = c(8, 10, 12, 14, 16, 18),
      labels = c("08:00","10:00","12:00",
                 "14:00","16:00","18:00")
    ) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = expansion(mult = c(0.02, 0.10))
    ) +
    labs(
      title    = sprintf(
        "Fig. 1%s \u2014 Predicted probability of behaviour across time of day (%s %s)",
        fig_letter, age, sex),
      subtitle = paste0(
        "Age = ", age, " \u00b7 Sex = ", sex,
        " \u00b7 solid = LASSO prediction",
        " \u00b7 dashed = bootstrap 95% CI (n=", N_BOOT, ")",
        " \u00b7 circles = observed proportions (n \u2265 2)"),
      x       = "Hour of day",
      y       = "Probability of Behaviour",
      caption = paste0(
        "Unweighted LASSO \u00b7 \u03bb = lambda.min",
        " \u00b7 Green = Forest",
        " \u00b7 Red = Roadside",
        " \u00b7 Purple = RoadsideVeg")
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.background  = element_rect(fill = "white", color = NA),
      panel.background = element_rect(fill = "white", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
      strip.text       = element_text(size = 10, face = "bold",
                                      color = "grey15"),
      strip.background = element_rect(fill = "grey97", color = "grey88"),
      axis.text.x      = element_text(angle = 30, hjust = 1, size = 8.5),
      axis.text.y      = element_text(size = 9),
      legend.position  = "top",
      legend.title     = element_text(size = 9.5, face = "bold"),
      legend.text      = element_text(size = 9.5),
      plot.title       = element_text(size = 11, face = "bold",
                                      color = "grey10"),
      plot.subtitle    = element_text(size = 8.5, color = "grey45"),
      plot.caption     = element_text(size = 7.5, color = "grey55",
                                      hjust = 0),
      plot.margin      = unit(c(10, 10, 8, 10), "pt")
    )

  fname <- sprintf("fig1%s_%s_%s_boot.png",
                   fig_letter,
                   tolower(age),
                   tolower(sex))
  ggsave(fname, plot = p, width = 15, height = 6,
         dpi = 300, bg = "white")
  cat(sprintf("Saved: %s\n", fname))
  p
}

# ── Run all 4 figures ─────────────────────────────────────────────
# Set N_BOOT <- 50 at top for quick test first
# Then change to 500 for final run
fig1a_boot <- make_fig1_boot("Adult",    "Female", "fig1a")
fig1b_boot <- make_fig1_boot("Adult",    "Male",   "fig1b")
fig1c_boot <- make_fig1_boot("Subadult", "Female", "fig1c")
fig1d_boot <- make_fig1_boot("Subadult", "Male",   "fig1d")

cat("\nAll Fig 1 bootstrap plots complete.\n")

```
#BOOTSTRAP PP


```{r}
# ══════════════════════════════════════════════════════════════════
# FIG 2a-2d — Predicted probability across time of day
# WITH Bootstrap 95% CI — SHADED RIBBON style
# Hours: 08:00–18:00
# All 4 Age×Sex combinations
# Requires: lasso_final_80, train_80, y_train_80
#
# SPEED SETTINGS:
#   Quick test:  N_BOOT <- 50,  NFOLDS_BOOT <- 3
#   Final run:   N_BOOT <- 500, NFOLDS_BOOT <- 5
# ══════════════════════════════════════════════════════════════════

library(ggplot2)
library(dplyr)
library(tidyr)
library(glmnet)

margin <- ggplot2::margin

# ── SPEED SETTINGS ────────────────────────────────────────────────
N_BOOT      <- 50    # change to 500 for final run
NFOLDS_BOOT <- 3     # change to 5 for final run

set.seed(123)

hours_fine <- seq(8, 18, by = 0.5)
beh_levels <- levels(y_train_80)

loc_colors <- c(
  "Forest"      = "#1D9E75",
  "Roadside"    = "#E24B4A",
  "RoadsideVeg" = "#7F77DD"
)

# ── Extract fitted model once ─────────────────────────────────────
lasso_fit   <- lasso_final_80$glmnet.fit
best_lambda <- lasso_final_80$lambda.min

# ── Build prediction grid ─────────────────────────────────────────
make_time_grid <- function(age, sex, loc) {
  data.frame(
    Location_Coarse = factor(loc,
                             levels = c("Forest","Roadside","RoadsideVeg")),
    Age.Class       = factor(age,  levels = levels(train_80$Age.Class)),
    Sex             = factor(sex,  levels = levels(train_80$Sex)),
    hour            = hours_fine,
    hour_sin        = sin(2 * pi * hours_fine / 24),
    hour_cos        = cos(2 * pi * hours_fine / 24)
  )
}

# ── Get LASSO predicted probabilities ─────────────────────────────
get_probs <- function(grid, fit = lasso_fit, lam = best_lambda) {
  X_new <- model.matrix(
    ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
    data = grid
  )[, -1]
  prob <- predict(fit,
                  newx = X_new,
                  s    = lam,
                  type = "response")[,,1]
  colnames(prob) <- beh_levels
  df          <- as.data.frame(prob)
  df$Hour     <- grid$hour
  df$Location <- as.character(grid$Location_Coarse[1])
  df
}

# ── Point estimates for all 3 locations ───────────────────────────
get_point_estimates <- function(age, sex) {
  bind_rows(lapply(
    c("Forest", "Roadside", "RoadsideVeg"),
    function(loc) get_probs(make_time_grid(age, sex, loc))
  ))
}

# ── Observed proportions (circles) ────────────────────────────────
obs_props <- function(age, sex) {
  totals <- train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour) |>
    summarise(n_total = n(), .groups = "drop")

  train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour, Behaviour_Coarse) |>
    summarise(n_beh = n(), .groups = "drop") |>
    left_join(totals, by = c("Location_Coarse", "hour")) |>
    mutate(prop = n_beh / n_total) |>
    filter(n_total >= 2) |>
    rename(Behaviour = Behaviour_Coarse,
           Hour      = hour,
           Location  = Location_Coarse)
}

# ── Bootstrap CI ──────────────────────────────────────────────────
# Refits LASSO on each bootstrap resample of full training data
# Factor levels re-enforced inside loop — prevents Roadside/RoadsideVeg swap
# NFOLDS_BOOT controls speed — use 3 for testing, 5 for final
boot_ci_time <- function(age, sex, n_boot = N_BOOT) {

  locs    <- c("Forest", "Roadside", "RoadsideVeg")
  results <- list()

  for (loc in locs) {

    sub   <- train_80 |>
      filter(Age.Class       == age,
             Sex             == sex,
             Location_Coarse == loc)

    n_sub <- nrow(sub)

    if (n_sub < 3) {
      cat(sprintf("  Skipping %s %s %s — n=%d too small\n",
                  age, sex, loc, n_sub))
      next
    }

    n_hours  <- length(hours_fine)
    boot_mat <- array(NA,
                      dim      = c(n_boot, n_hours, length(beh_levels)),
                      dimnames = list(NULL, NULL, beh_levels))
    n_success <- 0

    for (b in seq_len(n_boot)) {

      # Resample full training set with replacement
      idx_boot   <- sample(seq_len(nrow(train_80)), replace = TRUE)
      train_boot <- train_80[idx_boot, ]

      # CRITICAL: re-enforce correct factor levels after resampling
      # Without this, resampling resets to alphabetical order
      # which swaps Roadside and RoadsideVeg predictions
      train_boot$Location_Coarse <- factor(
        train_boot$Location_Coarse,
        levels = c("Forest", "Roadside", "RoadsideVeg"))
      train_boot$Age.Class <- factor(
        train_boot$Age.Class,
        levels = levels(train_80$Age.Class))
      train_boot$Sex <- factor(
        train_boot$Sex,
        levels = levels(train_80$Sex))
      train_boot$Behaviour_Coarse <- factor(
        train_boot$Behaviour_Coarse,
        levels = levels(train_80$Behaviour_Coarse))

      # Rebuild model matrix from bootstrap sample
      X_boot <- tryCatch(
        model.matrix(
          ~ (Location_Coarse + Age.Class + Sex +
               hour_sin + hour_cos)^3,
          data = train_boot)[, -1],
        error = function(e) NULL)
      if (is.null(X_boot)) next

      y_boot <- train_boot$Behaviour_Coarse

      # Refit LASSO — use NFOLDS_BOOT for speed control
      fit_boot <- tryCatch(
        cv.glmnet(
          x      = X_boot,
          y      = y_boot,
          family = "multinomial",
          alpha  = 1,
          nfolds = NFOLDS_BOOT),   # 3 = faster, 5 = more accurate
        error = function(e) NULL)
      if (is.null(fit_boot)) next

      # Predict on fixed grid
      grid <- make_time_grid(age, sex, loc)
      prob <- tryCatch(
        get_probs(grid,
                  fit = fit_boot$glmnet.fit,
                  lam = fit_boot$lambda.min),
        error = function(e) NULL)
      if (is.null(prob)) next

      boot_mat[b, , ] <- as.matrix(prob[, beh_levels])
      n_success <- n_success + 1

      # Progress every 10 boots
      if (b %% 10 == 0) {
        cat(sprintf("    %s %s %s — boot %d/%d\n",
                    age, sex, loc, b, n_boot))
      }
    }

    cat(sprintf("  Done: %s %s %s (n=%d, %d/%d succeeded)\n",
                age, sex, loc, n_sub, n_success, n_boot))

    # Compute 2.5th and 97.5th percentiles
    ci_list <- lapply(seq_along(beh_levels), function(k) {
      vals <- boot_mat[, , k]
      vals <- vals[complete.cases(vals), , drop = FALSE]
      data.frame(
        Hour      = hours_fine,
        Behaviour = beh_levels[k],
        Lower     = apply(vals, 2, quantile, 0.025, na.rm = TRUE),
        Upper     = apply(vals, 2, quantile, 0.975, na.rm = TRUE),
        Location  = loc)
    })

    results[[loc]] <- bind_rows(ci_list)
  }

  bind_rows(results)
}

# ── Main plot function ─────────────────────────────────────────────
make_fig2_boot <- function(age, sex, fig_label) {

  cat(sprintf("\n══════════════════════════════════════\n"))
  cat(sprintf("Building %s (%s %s)...\n", fig_label, age, sex))
  cat(sprintf("N_BOOT=%d  NFOLDS=%d\n", N_BOOT, NFOLDS_BOOT))
  cat(sprintf("══════════════════════════════════════\n"))

  fig_letter <- substr(fig_label, nchar(fig_label), nchar(fig_label))

  # Point estimates
  point_df <- get_point_estimates(age, sex) |>
    pivot_longer(cols      = all_of(beh_levels),
                 names_to  = "Behaviour",
                 values_to = "Prob") |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # Bootstrap CIs
  ci_df <- boot_ci_time(age, sex) |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # Merge CI with point estimates for ribbon
  ribbon_df <- ci_df |>
    left_join(point_df, by = c("Hour", "Behaviour", "Location"))

  # Observed proportions
  obs_df <- obs_props(age, sex) |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # ── Plot ──────────────────────────────────────────────────────────
  p <- ggplot() +

    # Shaded bootstrap CI ribbon
    geom_ribbon(
      data  = ribbon_df,
      aes(x    = Hour,
          ymin = Lower,
          ymax = Upper,
          fill = Location),
      alpha = 0.18
    ) +

    # LASSO point estimate — solid line
    geom_line(
      data      = point_df,
      aes(x = Hour, y = Prob, color = Location),
      linewidth = 1.1
    ) +

    # Observed proportion circles
    geom_point(
      data   = obs_df,
      aes(x = Hour, y = prop, color = Location),
      size   = 2.5,
      shape  = 21,
      fill   = "white",
      stroke = 1.2,
      alpha  = 0.85
    ) +

    facet_wrap(~ Behaviour, nrow = 1, scales = "free_y") +
    scale_color_manual(values = loc_colors, name = "Location") +
    scale_fill_manual(values  = loc_colors, name = "Location") +
    scale_x_continuous(
      limits = c(8, 18),
      breaks = c(8, 10, 12, 14, 16, 18),
      labels = c("08:00","10:00","12:00",
                 "14:00","16:00","18:00")
    ) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = expansion(mult = c(0.02, 0.10))
    ) +
    labs(
      title    = sprintf(
        "Fig. 2%s \u2014 Predicted probability of behaviour across time of day (%s %s)",
        fig_letter, age, sex),
      subtitle = paste0(
        "Age = ", age, " \u00b7 Sex = ", sex,
        " \u00b7 solid = LASSO prediction",
        " \u00b7 shading = bootstrap 95% CI (n=", N_BOOT, ")",
        " \u00b7 circles = observed proportions (n \u2265 2)"),
      x       = "Hour of day",
      y       = "Probability of Behaviour",
      caption = paste0(
        "Unweighted LASSO \u00b7 \u03bb = lambda.min",
        " \u00b7 Green = Forest",
        " \u00b7 Red = Roadside",
        " \u00b7 Purple = RoadsideVeg")
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.background  = element_rect(fill = "white", color = NA),
      panel.background = element_rect(fill = "white", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
      strip.text       = element_text(size = 10, face = "bold",
                                      color = "grey15"),
      strip.background = element_rect(fill = "grey97", color = "grey88"),
      axis.text.x      = element_text(angle = 30, hjust = 1, size = 8.5),
      axis.text.y      = element_text(size = 9),
      legend.position  = "top",
      legend.title     = element_text(size = 9.5, face = "bold"),
      legend.text      = element_text(size = 9.5),
      plot.title       = element_text(size = 11, face = "bold",
                                      color = "grey10"),
      plot.subtitle    = element_text(size = 8.5, color = "grey45"),
      plot.caption     = element_text(size = 7.5, color = "grey55",
                                      hjust = 0),
      plot.margin      = unit(c(10, 10, 8, 10), "pt")
    )

  # Save
  fname <- sprintf("fig2%s_%s_%s_boot.png",
                   fig_letter,
                   tolower(age),
                   tolower(sex))
  ggsave(fname, plot = p, width = 15, height = 6,
         dpi = 300, bg = "white")
  cat(sprintf("Saved: %s\n", fname))
  p
}

# ══════════════════════════════════════════════════════════════════
# RUN — test one figure first then all four
# ══════════════════════════════════════════════════════════════════

# ── STEP 1: Test with fig2a first (fastest subgroup) ─────────────
cat("Running quick test: fig2a (Adult Female)\n")
cat("N_BOOT =", N_BOOT, "— change to 500 for final run\n\n")

fig2a_boot <- make_fig2_boot("Adult", "Female", "fig2a")

# ── STEP 2: If fig2a looks correct run all four ───────────────────
# Uncomment after confirming fig2a is correct:

```

```{r}

fig2b_boot <- make_fig2_boot("Adult",    "Male",   "fig2b")
```
```{r}
 fig2c_boot <- make_fig2_boot("Subadult", "Female", "fig2c")
```
```{r}
 fig2d_boot <- make_fig2_boot("Subadult", "Male",   "fig2d")

```

# cat("\nAll Fig 2 bootstrap plots complete.\n")
# cat("Files saved:\n")
# cat("  fig2a_adult_female_boot.png\n")
# cat("  fig2b_adult_male_boot.png\n")
# cat("  fig2c_subadult_female_boot.png\n")
# cat("  fig2d_subadult_male_boot.png\n")


#Again Bootstrap for axis improvement

```{r}
# ══════════════════════════════════════════════════════════════════
# FIG 1a-1d — Predicted probability across time of day
# WITH Bootstrap 95% CI — SHADED RIBBON style
# Hours: 08:00–18:00
# All 4 Age×Sex combinations
# Requires: lasso_final_80, train_80, y_train_80
#
# SPEED SETTINGS:
#   Quick test:  N_BOOT <- 50,  NFOLDS_BOOT <- 3
#   Final run:   N_BOOT <- 500, NFOLDS_BOOT <- 5
# ══════════════════════════════════════════════════════════════════

library(ggplot2)
library(dplyr)
library(tidyr)
library(glmnet)

margin <- ggplot2::margin

# ── SPEED SETTINGS ────────────────────────────────────────────────
N_BOOT      <- 50    # change to 500 for final run
NFOLDS_BOOT <- 3     # change to 5 for final run

set.seed(123)

hours_fine <- seq(8, 18, by = 0.5)
beh_levels <- levels(y_train_80)

loc_colors <- c(
  "Forest"      = "#1D9E75",
  "Roadside"    = "#E24B4A",
  "RoadsideVeg" = "#7F77DD"
)

# ── Extract fitted model once ─────────────────────────────────────
lasso_fit   <- lasso_final_80$glmnet.fit
best_lambda <- lasso_final_80$lambda.min

# ── Build prediction grid ─────────────────────────────────────────
make_time_grid <- function(age, sex, loc) {
  data.frame(
    Location_Coarse = factor(loc,
                             levels = c("Forest","Roadside","RoadsideVeg")),
    Age.Class       = factor(age,  levels = levels(train_80$Age.Class)),
    Sex             = factor(sex,  levels = levels(train_80$Sex)),
    hour            = hours_fine,
    hour_sin        = sin(2 * pi * hours_fine / 24),
    hour_cos        = cos(2 * pi * hours_fine / 24)
  )
}

# ── Get LASSO predicted probabilities ─────────────────────────────
get_probs <- function(grid, fit = lasso_fit, lam = best_lambda) {
  X_new <- model.matrix(
    ~ (Location_Coarse + Age.Class + Sex + hour_sin + hour_cos)^3,
    data = grid
  )[, -1]
  prob <- predict(fit,
                  newx = X_new,
                  s    = lam,
                  type = "response")[,,1]
  colnames(prob) <- beh_levels
  df          <- as.data.frame(prob)
  df$Hour     <- grid$hour
  df$Location <- as.character(grid$Location_Coarse[1])
  df
}

# ── Point estimates for all 3 locations ───────────────────────────
get_point_estimates <- function(age, sex) {
  bind_rows(lapply(
    c("Forest", "Roadside", "RoadsideVeg"),
    function(loc) get_probs(make_time_grid(age, sex, loc))
  ))
}

# ── Observed proportions (circles) ────────────────────────────────
obs_props <- function(age, sex) {
  totals <- train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour) |>
    summarise(n_total = n(), .groups = "drop")

  train_80 |>
    filter(Age.Class == age, Sex == sex,
           hour >= 8, hour <= 18) |>
    group_by(Location_Coarse, hour, Behaviour_Coarse) |>
    summarise(n_beh = n(), .groups = "drop") |>
    left_join(totals, by = c("Location_Coarse", "hour")) |>
    mutate(prop = n_beh / n_total) |>
    filter(n_total >= 2) |>
    rename(Behaviour = Behaviour_Coarse,
           Hour      = hour,
           Location  = Location_Coarse)
}

# ── Bootstrap CI ──────────────────────────────────────────────────
# Refits LASSO on each bootstrap resample of full training data
# Factor levels re-enforced inside loop — prevents Roadside/RoadsideVeg swap
# NFOLDS_BOOT controls speed — use 3 for testing, 5 for final
boot_ci_time <- function(age, sex, n_boot = N_BOOT) {

  locs    <- c("Forest", "Roadside", "RoadsideVeg")
  results <- list()

  for (loc in locs) {

    sub   <- train_80 |>
      filter(Age.Class       == age,
             Sex             == sex,
             Location_Coarse == loc)

    n_sub <- nrow(sub)

    if (n_sub < 3) {
      cat(sprintf("  Skipping %s %s %s — n=%d too small\n",
                  age, sex, loc, n_sub))
      next
    }

    n_hours  <- length(hours_fine)
    boot_mat <- array(NA,
                      dim      = c(n_boot, n_hours, length(beh_levels)),
                      dimnames = list(NULL, NULL, beh_levels))
    n_success <- 0

    for (b in seq_len(n_boot)) {

      # Resample full training set with replacement
      idx_boot   <- sample(seq_len(nrow(train_80)), replace = TRUE)
      train_boot <- train_80[idx_boot, ]

      # CRITICAL: re-enforce correct factor levels after resampling
      # Without this, resampling resets to alphabetical order
      # which swaps Roadside and RoadsideVeg predictions
      train_boot$Location_Coarse <- factor(
        train_boot$Location_Coarse,
        levels = c("Forest", "Roadside", "RoadsideVeg"))
      train_boot$Age.Class <- factor(
        train_boot$Age.Class,
        levels = levels(train_80$Age.Class))
      train_boot$Sex <- factor(
        train_boot$Sex,
        levels = levels(train_80$Sex))
      train_boot$Behaviour_Coarse <- factor(
        train_boot$Behaviour_Coarse,
        levels = levels(train_80$Behaviour_Coarse))

      # Rebuild model matrix from bootstrap sample
      X_boot <- tryCatch(
        model.matrix(
          ~ (Location_Coarse + Age.Class + Sex +
               hour_sin + hour_cos)^3,
          data = train_boot)[, -1],
        error = function(e) NULL)
      if (is.null(X_boot)) next

      y_boot <- train_boot$Behaviour_Coarse

      # Refit LASSO — use NFOLDS_BOOT for speed control
      fit_boot <- tryCatch(
        cv.glmnet(
          x      = X_boot,
          y      = y_boot,
          family = "multinomial",
          alpha  = 1,
          nfolds = NFOLDS_BOOT),   # 3 = faster, 5 = more accurate
        error = function(e) NULL)
      if (is.null(fit_boot)) next

      # Predict on fixed grid
      grid <- make_time_grid(age, sex, loc)
      prob <- tryCatch(
        get_probs(grid,
                  fit = fit_boot$glmnet.fit,
                  lam = fit_boot$lambda.min),
        error = function(e) NULL)
      if (is.null(prob)) next

      boot_mat[b, , ] <- as.matrix(prob[, beh_levels])
      n_success <- n_success + 1

      # Progress every 10 boots
      if (b %% 10 == 0) {
        cat(sprintf("    %s %s %s — boot %d/%d\n",
                    age, sex, loc, b, n_boot))
      }
    }

    cat(sprintf("  Done: %s %s %s (n=%d, %d/%d succeeded)\n",
                age, sex, loc, n_sub, n_success, n_boot))

    # Compute 2.5th and 97.5th percentiles
    ci_list <- lapply(seq_along(beh_levels), function(k) {
      vals <- boot_mat[, , k]
      vals <- vals[complete.cases(vals), , drop = FALSE]
      data.frame(
        Hour      = hours_fine,
        Behaviour = beh_levels[k],
        Lower     = apply(vals, 2, quantile, 0.025, na.rm = TRUE),
        Upper     = apply(vals, 2, quantile, 0.975, na.rm = TRUE),
        Location  = loc)
    })

    results[[loc]] <- bind_rows(ci_list)
  }

  bind_rows(results)
}

# ── Main plot function ─────────────────────────────────────────────
make_fig1_boot <- function(age, sex, fig_label) {

  cat(sprintf("\n══════════════════════════════════════\n"))
  cat(sprintf("Building %s (%s %s)...\n", fig_label, age, sex))
  cat(sprintf("N_BOOT=%d  NFOLDS=%d\n", N_BOOT, NFOLDS_BOOT))
  cat(sprintf("══════════════════════════════════════\n"))

  fig_letter <- substr(fig_label, nchar(fig_label), nchar(fig_label))

  # Point estimates
  point_df <- get_point_estimates(age, sex) |>
    pivot_longer(cols      = all_of(beh_levels),
                 names_to  = "Behaviour",
                 values_to = "Prob") |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # Bootstrap CIs
  ci_df <- boot_ci_time(age, sex) |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # Merge CI with point estimates for ribbon
  ribbon_df <- ci_df |>
    left_join(point_df, by = c("Hour", "Behaviour", "Location"))

  # Observed proportions
  obs_df <- obs_props(age, sex) |>
    mutate(
      Behaviour = factor(Behaviour, levels = beh_levels),
      Location  = factor(Location,
                         levels = c("Forest","Roadside","RoadsideVeg"))
    )

  # ── Plot ──────────────────────────────────────────────────────────
  p <- ggplot() +

    # Shaded bootstrap CI ribbon
    geom_ribbon(
      data  = ribbon_df,
      aes(x    = Hour,
          ymin = Lower,
          ymax = Upper,
          fill = Location),
      alpha = 0.18
    ) +

    # LASSO point estimate — solid line
    geom_line(
      data      = point_df,
      aes(x = Hour, y = Prob, color = Location),
      linewidth = 1.1
    ) +

    # Observed proportion circles
    geom_point(
      data   = obs_df,
      aes(x = Hour, y = prop, color = Location),
      size   = 2.5,
      shape  = 21,
      fill   = "white",
      stroke = 1.2,
      alpha  = 0.85
    ) +

    # CHANGED: removed scales = "free_y" so all panels share one fixed scale
    facet_wrap(~ Behaviour, nrow = 1) +
    scale_color_manual(values = loc_colors, name = "Location") +
    scale_fill_manual(values  = loc_colors, name = "Location") +
    scale_x_continuous(
      limits = c(8, 18),
      breaks = c(8, 10, 12, 14, 16, 18),
      labels = c("08:00","10:00","12:00",
                 "14:00","16:00","18:00")
    ) +
    # CHANGED: fixed 0–100% range so every panel & every figure is comparable
    scale_y_continuous(
      limits = c(0, 1),
      labels = scales::percent_format(accuracy = 1),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    labs(
      title    = sprintf(
        "Fig. 1%s \u2014 Predicted probability of behaviour across time of day (%s %s)",
        fig_letter, age, sex),
      subtitle = paste0(
        "Age = ", age, " \u00b7 Sex = ", sex,
        " \u00b7 solid = LASSO prediction",
        " \u00b7 shading = bootstrap 95% CI (n=", N_BOOT, ")",
        " \u00b7 circles = observed proportions (n \u2265 2)"),
      x       = "Hour of day",
      y       = "Probability of Behaviour",
      caption = paste0(
        "Unweighted LASSO \u00b7 \u03bb = lambda.min",
        " \u00b7 Green = Forest",
        " \u00b7 Red = Roadside",
        " \u00b7 Purple = RoadsideVeg")
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.background  = element_rect(fill = "white", color = NA),
      panel.background = element_rect(fill = "white", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
      strip.text       = element_text(size = 10, face = "bold",
                                      color = "grey15"),
      strip.background = element_rect(fill = "grey97", color = "grey88"),
      axis.text.x      = element_text(angle = 30, hjust = 1, size = 8.5),
      axis.text.y      = element_text(size = 9),
      legend.position  = "top",
      legend.title     = element_text(size = 9.5, face = "bold"),
      legend.text      = element_text(size = 9.5),
      plot.title       = element_text(size = 11, face = "bold",
                                      color = "grey10"),
      plot.subtitle    = element_text(size = 8.5, color = "grey45"),
      plot.caption     = element_text(size = 7.5, color = "grey55",
                                      hjust = 0),
      plot.margin      = unit(c(10, 10, 8, 10), "pt")
    )

  # Save
  fname <- sprintf("fig1%s_%s_%s_boot.png",
                   fig_letter,
                   tolower(age),
                   tolower(sex))
  ggsave(fname, plot = p, width = 15, height = 6,
         dpi = 300, bg = "white")
  cat(sprintf("Saved: %s\n", fname))
  p
}

# ══════════════════════════════════════════════════════════════════
# RUN — test one figure first then all four
# ══════════════════════════════════════════════════════════════════

# ── STEP 1: Test with fig1a first (fastest subgroup) ─────────────
cat("Running quick test: fig1a (Adult Female)\n")
cat("N_BOOT =", N_BOOT, "— change to 500 for final run\n\n")

fig1a_boot <- make_fig1_boot("Adult", "Female", "fig1a")

```

```{r}
# ── STEP 2: If fig1a looks correct run all four ───────────────────
# Uncomment after confirming fig1a is correct:

fig1b_boot <- make_fig1_boot("Adult",    "Male",   "fig1b")

```

```{r}
fig1c_boot <- make_fig1_boot("Subadult", "Female", "fig1c")

```

```{r}
fig1d_boot <- make_fig1_boot("Subadult", "Male",   "fig1d")

```

# cat("\nAll Fig 1 bootstrap plots complete.\n")
# cat("Files saved:\n")
# cat("  fig1a_adult_female_boot.png\n")
# cat("  fig1b_adult_male_boot.png\n")
# cat("  fig1c_subadult_female_boot.png\n")
# cat("  fig1d_subadult_male_boot.png\n")
```

```{r}
# ══════════════════════════════════════════════════════════════════
# BOOTSTRAP PREDICTIONS — GROUPED BAR CHART WITH 95% CI ERROR BARS
# ══════════════════════════════════════════════════════════════════

library(ggplot2)
library(dplyr)

# Load data
df <- read.csv("fig2_bootstrap_predictions.csv")

# Clean up factor levels
df$Location <- factor(df$Location, levels = c("Forest", "Roadside", "RoadsideVeg"))
df$Behaviour <- factor(df$Behaviour, levels = c("Rest", "Feed/Forage", 
                                                  "Human Directed", "Move", 
                                                  "Socio-Sexual"))
df$Combination <- factor(df$Combination, 
                         levels = c("Adult Female", "Adult Male", 
                                    "Subadult Female", "Subadult Male"))

# Convert to percentage
df <- df %>%
  mutate(
    Mean_pct    = Mean     * 100,
    Lower_pct   = Lower95  * 100,
    Upper_pct   = Upper95  * 100
  )

# Colour palette matching your existing figures
loc_colours <- c(
  "Forest"      = "#2ecc71",
  "Roadside"    = "#e74c3c",
  "RoadsideVeg" = "#9b59b6"
)

# ── PLOT: all behaviours × all groups ─────────────────────────────
p <- ggplot(df, aes(x = Location, y = Mean_pct, fill = Location)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(
    aes(ymin = Lower_pct, ymax = Upper_pct),
    width = 0.25,
    linewidth = 0.6,
    colour = "grey30"
  ) +
  facet_grid(Combination ~ Behaviour) +
  scale_fill_manual(values = loc_colours) +
  scale_y_continuous(
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    title   = "Predicted probability of behaviour by location and demographic group",
    subtitle = "Bootstrap mean ± 95% CI · LASSO (λ = λ.min)",
    x       = NULL,
    y       = "Probability of Behaviour (%)",
    fill    = "Location",
    caption = "Rows = Age × Sex group · Columns = Behaviour category"
  ) +
  theme_bw(base_size = 11) +
  theme(
    strip.background   = element_rect(fill = "grey92", colour = "grey70"),
    strip.text         = element_text(face = "bold", size = 9),
    axis.text.x        = element_blank(),
    axis.ticks.x       = element_blank(),
    legend.position    = "top",
    legend.title       = element_text(face = "bold"),
    panel.grid.major.x = element_blank(),
    plot.title         = element_text(face = "bold", size = 12),
    plot.subtitle      = element_text(size = 9, colour = "grey40"),
    plot.caption       = element_text(size = 8, colour = "grey50")
  )

print(p)

# Save
ggsave("fig3_bootstrap_bar_chart.png", plot = p,
       width = 14, height = 10, dpi = 300)
cat("Saved: fig3_bootstrap_bar_chart.png\n")

# ── PLOT 2: Human Directed only — cleaner focus ────────────────────
hd_only <- df %>% filter(Behaviour == "Human Directed")

p_hd <- ggplot(hd_only, aes(x = Combination, y = Mean_pct, fill = Location)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.75),
           width = 0.65) +
  geom_errorbar(
    aes(ymin = Lower_pct, ymax = Upper_pct),
    position = position_dodge(width = 0.75),
    width = 0.25,
    linewidth = 0.6,
    colour = "grey30"
  ) +
  geom_text(
    aes(label = paste0(round(Mean_pct, 1), "%"), y = Upper_pct + 1),
    position = position_dodge(width = 0.75),
    size = 3, vjust = 0
  ) +
  scale_fill_manual(values = loc_colours) +
  scale_y_continuous(
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title    = "Predicted Human Directed probability by location and demographic group",
    subtitle = "Bootstrap mean ± 95% CI · Unweighted LASSO (λ = λ.min)",
    x        = "Age × Sex Group",
    y        = "Probability of Human Directed Behaviour (%)",
    fill     = "Location"
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position    = "top",
    legend.title       = element_text(face = "bold"),
    panel.grid.major.x = element_blank(),
    plot.title         = element_text(face = "bold", size = 12),
    plot.subtitle      = element_text(size = 9, colour = "grey40")
  )

print(p_hd)

ggsave("fig3b_HD_bootstrap_bar.png", plot = p_hd,
       width = 10, height = 6, dpi = 300)
cat("Saved: fig3b_HD_bootstrap_bar.png\n")
```

#K=2 (only)

```{r}
# ══════════════════════════════════════════════════════════════════
# REFIT LASSO — k=2 ONLY (no k=1)
# hour_sin2 and hour_cos2 only — no hour_sin or hour_cos
# ══════════════════════════════════════════════════════════════════

# ── STEP 1: Add second harmonic ───────────────────────────────────
train_80$hour_sin2 <- sin(4 * pi * train_80$hour / 24)
train_80$hour_cos2 <- cos(4 * pi * train_80$hour / 24)

test_20$hour_sin2  <- sin(4 * pi * test_20$hour / 24)
test_20$hour_cos2  <- cos(4 * pi * test_20$hour / 24)
```

```{r}

# ── STEP 2: Build model matrix — k=2 only ────────────────────────
X_train_k2only <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex +
     hour_sin2 + hour_cos2)^3,
  data = train_80
)[, -1]

X_test_k2only <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex +
     hour_sin2 + hour_cos2)^3,
  data = test_20
)[, -1]

cat("Model matrix dimensions (k=2 only):\n")
cat("Train:", nrow(X_train_k2only), "x", ncol(X_train_k2only), "\n")
cat("Test: ", nrow(X_test_k2only),  "x", ncol(X_test_k2only),  "\n\n")
```

```{r}

# ── STEP 3: Refit LASSO ───────────────────────────────────────────
set.seed(123)
lasso_k2only <- cv.glmnet(
  x      = X_train_k2only,
  y      = y_train_80,
  family = "multinomial",
  alpha  = 1,
  nfolds = 10
)

cat("LASSO k=2 only fitted.\n")
cat("lambda.min =", lasso_k2only$lambda.min, "\n\n")
```

```{r}

# ── STEP 4: Test set predictions ─────────────────────────────────
prob_k2only_test <- predict(lasso_k2only,
                             newx = X_test_k2only,
                             s    = "lambda.min",
                             type = "response")[,,1]
colnames(prob_k2only_test) <- levels(y_test_20)

pred_k2only_test <- predict(lasso_k2only,
                             newx = X_test_k2only,
                             s    = "lambda.min",
                             type = "class")[, 1]

y_onehot_test <- model.matrix(
  ~ y_test_20 - 1,
  data = data.frame(y_test_20 = y_test_20))
colnames(y_onehot_test) <- levels(y_test_20)
```

```{r}

# ── STEP 5: Compare k=1 vs k=2 only ──────────────────────────────
loss_k1     <- compute_01_loss(
  predict(lasso_final_80, newx = X_test_20,
          s = "lambda.min", type = "class")[,1],
  y_test_20)
loss_k2only <- compute_01_loss(pred_k2only_test, y_test_20)

ll_k1     <- compute_logloss(
  predict(lasso_final_80, newx = X_test_20,
          s = "lambda.min", type = "response")[,,1],
  y_onehot_test)
ll_k2only <- compute_logloss(prob_k2only_test, y_onehot_test)

br_k1     <- compute_brier(
  predict(lasso_final_80, newx = X_test_20,
          s = "lambda.min", type = "response")[,,1],
  y_onehot_test)
br_k2only <- compute_brier(prob_k2only_test, y_onehot_test)

cat("══════════════════════════════════════════\n")
cat("LASSO k=1 vs k=2 only — Test Set\n")
cat("══════════════════════════════════════════\n\n")
cat(sprintf("%-15s  %8s  %10s\n", "Metric", "k=1", "k=2 only"))
cat(strrep("─", 38), "\n")
cat(sprintf("%-15s  %8.4f  %10.4f\n", "0-1 Loss ↓",  loss_k1, loss_k2only))
cat(sprintf("%-15s  %8.4f  %10.4f\n", "Log-Loss ↓",  ll_k1,   ll_k2only))
cat(sprintf("%-15s  %8.4f  %10.4f\n", "Brier ↓",     br_k1,   br_k2only))
cat(strrep("─", 38), "\n\n")
```

```{r}

# ── STEP 6: HD sanity check ───────────────────────────────────────
hour_fixed <- 8
test_locs <- data.frame(
  Location_Coarse = factor(
    c("Forest", "Roadside", "RoadsideVeg"),
    levels = c("Forest", "Roadside", "RoadsideVeg")),
  Age.Class  = factor("Adult",  levels = levels(train_80$Age.Class)),
  Sex        = factor("Female", levels = levels(train_80$Sex)),
  hour       = hour_fixed,
  hour_sin2  = sin(4 * pi * hour_fixed / 24),
  hour_cos2  = cos(4 * pi * hour_fixed / 24)
)

X_check <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex +
     hour_sin2 + hour_cos2)^3,
  data = test_locs
)[, -1]

prob_check <- predict(lasso_k2only,
                      newx = X_check,
                      s    = "lambda.min",
                      type = "response")[,,1]
colnames(prob_check) <- levels(y_train_80)
rownames(prob_check) <- c("Forest", "Roadside", "RoadsideVeg")

cat("HD probabilities k=2 only (Roadside must be highest):\n")
print(round(prob_check[, "Human Directed", drop = FALSE], 3))
```
#(K=2+k=1) (circular time change)

```{r}
# ══════════════════════════════════════════════════════════════════
# REFIT LASSO — k=2 Fourier harmonics (unweighted)
# Adds second harmonic to capture bimodal daily patterns
# ══════════════════════════════════════════════════════════════════

# ── STEP 1: Add second harmonic to train and test ─────────────────
train_80$hour_sin2 <- sin(4 * pi * train_80$hour / 24)
train_80$hour_cos2 <- cos(4 * pi * train_80$hour / 24)

test_20$hour_sin2  <- sin(4 * pi * test_20$hour / 24)
test_20$hour_cos2  <- cos(4 * pi * test_20$hour / 24)

cat("Second harmonic added to train and test.\n\n")
```


```{r}

# ── STEP 2: Rebuild model matrix with k=2 ────────────────────────
X_train_80_k2 <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex +
     hour_sin + hour_cos +
     hour_sin2 + hour_cos2)^3,
  data = train_80
)[, -1]

X_test_20_k2 <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex +
     hour_sin + hour_cos +
     hour_sin2 + hour_cos2)^3,
  data = test_20
)[, -1]

cat("Model matrix dimensions (k=2):\n")
cat("Train:", nrow(X_train_80_k2), "x", ncol(X_train_80_k2), "\n")
cat("Test: ", nrow(X_test_20_k2),  "x", ncol(X_test_20_k2),  "\n\n")
```

```{r}

# ── STEP 3: Refit LASSO on k=2 matrix ────────────────────────────
set.seed(123)
lasso_k2 <- cv.glmnet(
  x      = X_train_80_k2,
  y      = y_train_80,
  family = "multinomial",
  alpha  = 1,
  nfolds = 10
)

cat("LASSO k=2 fitted.\n")
cat("lambda.min =", lasso_k2$lambda.min, "\n")
cat("lambda.1se =", lasso_k2$lambda.1se, "\n\n")
```

```{r}

# ── STEP 4: Evaluate on test set ─────────────────────────────────
prob_lasso_k2_test <- predict(lasso_k2,
                               newx = X_test_20_k2,
                               s    = "lambda.min",
                               type = "response")[,,1]
colnames(prob_lasso_k2_test) <- levels(y_test_20)

pred_lasso_k2_test <- predict(lasso_k2,
                               newx = X_test_20_k2,
                               s    = "lambda.min",
                               type = "class")[, 1]

y_onehot_test_k2 <- model.matrix(
  ~ y_test_20 - 1,
  data = data.frame(y_test_20 = y_test_20))
colnames(y_onehot_test_k2) <- levels(y_test_20)
```

```{r}

# ── STEP 5: Compare k=1 vs k=2 metrics ───────────────────────────
loss_k1 <- compute_01_loss(
  predict(lasso_final_80, newx = X_test_20,
          s = "lambda.min", type = "class")[,1],
  y_test_20)

loss_k2 <- compute_01_loss(pred_lasso_k2_test, y_test_20)

ll_k1 <- compute_logloss(
  predict(lasso_final_80, newx = X_test_20,
          s = "lambda.min", type = "response")[,,1],
  y_onehot_test_k2)

ll_k2 <- compute_logloss(prob_lasso_k2_test, y_onehot_test_k2)

br_k1 <- compute_brier(
  predict(lasso_final_80, newx = X_test_20,
          s = "lambda.min", type = "response")[,,1],
  y_onehot_test_k2)

br_k2 <- compute_brier(prob_lasso_k2_test, y_onehot_test_k2)

cat("══════════════════════════════════════════\n")
cat("LASSO k=1 vs k=2 — Test Set Comparison\n")
cat("══════════════════════════════════════════\n\n")
cat(sprintf("%-15s  %8s  %10s\n", "Metric", "k=1", "k=1+k=2"))
cat(strrep("─", 35), "\n")
cat(sprintf("%-15s  %8.4f  %8.4f\n", "0-1 Loss ↓",  loss_k1, loss_k2))
cat(sprintf("%-15s  %8.4f  %8.4f\n", "Log-Loss ↓",  ll_k1,   ll_k2))
cat(sprintf("%-15s  %8.4f  %8.4f\n", "Brier ↓",     br_k1,   br_k2))
cat(strrep("─", 35), "\n\n")

```

```{r}

# ── STEP 6: Sanity check — HD probabilities ───────────────────────
hour_fixed <- 8
test_locs_k2 <- data.frame(
  Location_Coarse = factor(
    c("Forest", "Roadside", "RoadsideVeg"),
    levels = c("Forest", "Roadside", "RoadsideVeg")),
  Age.Class  = factor("Adult",  levels = levels(train_80$Age.Class)),
  Sex        = factor("Female", levels = levels(train_80$Sex)),
  hour       = hour_fixed,
  hour_sin   = sin(2 * pi * hour_fixed / 24),
  hour_cos   = cos(2 * pi * hour_fixed / 24),
  hour_sin2  = sin(4 * pi * hour_fixed / 24),
  hour_cos2  = cos(4 * pi * hour_fixed / 24)
)

X_check_k2 <- model.matrix(
  ~ (Location_Coarse + Age.Class + Sex +
     hour_sin + hour_cos +
     hour_sin2 + hour_cos2)^3,
  data = test_locs_k2
)[, -1]

prob_check_k2 <- predict(lasso_k2,
                          newx = X_check_k2,
                          s    = "lambda.min",
                          type = "response")[,,1]
colnames(prob_check_k2) <- levels(y_train_80)
rownames(prob_check_k2) <- c("Forest", "Roadside", "RoadsideVeg")

cat("HD probabilities k=2 (Roadside must be highest):\n")
print(round(prob_check_k2[, "Human Directed", drop = FALSE], 3))

```



```{r}
```


```{r}
```

```{r}
```


```{r}
```

```{r}
```

```{r}
```
